---
title: "Gapminder Analysis"
author: "Andrea Martinez"
date: "2024-02-07"
output: 
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide 
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the first line of my document. In this document, I will analyze how life expentency has changed over time

# Reading in gapminder data

```{r reading_in_data}
gapminder_df<-read.csv("data/gapminder_data.csv")

```

## A small header

### An even smaller header

##### A very small header

Lists:

-   List item 1
-   List item 2

Another type of list:

-   hyphen list
-   hyphen list

Numbered list:

1.  number one
2.  number two
    -   sub list
3.  number three
    -   sub list

We can make text **bold with two asterisks** or *italic with one asterisk*

We can also insert greek letters $\mu$ $\Delta$

We can make [hyperlinks](https://pad.carpentries.org/Spring2024-BIOMI6300)

# Calculating summary statistics

```{r calculating_median_pop, echo = FALSE}
med_pop <- median(gapminder_df$pop)

num_obs <- nrow(gapminder_df)
```

Over `r num_obs` observations, the median population was `r round(med_pop/10e6, 3)` million people

# Plotting life expectancy over the years

```{r line-graph, warning = FALSE}
library(ggplot2)

ggplot(gapminder_df, aes(x=year, y=lifeExp, group = country, color = continent)) +
         geom_line()+
  scale_y_log10() +
  scale_x_log10() +
  scale_x_continuous(limits = c(1997,2007))

# Make a line plot
# Log scale y axis
# Bonus: Change the theme
```

# Plotting Population Distribution

```{r histogram, message = FALSE, fig.width = 10}
ggplot(gapminder_df, aes(x=pop))+
  geom_histogram()
  
```

| Col1                | Col2 | Col3 |
|---------------------|------|------|
| put some stuff here |      |      |
|                     |      |      |
|                     |      |      |
